\chapter*{Abstract}

\section*{\tituloPortadaEngVal}
This bachelor's thesis presents the study, analysis, design, and implementation of a statistics and achievements page for the online judge \aer The main objective of this project is to enhance the user experience by providing a clear, structured, and intuitive visualization of their knowledge level and progress over time in the field of competitive programming. Additionally, it aims to provide an extra source of motivation through the incorporation of gamification techniques that encourage active participation and continuous learning. 

To define the system, a survey was conducted among students with experience or interest in competitive programming. The goal was to identify their needs, preferences, and perceived shortcomings in the current platform. Based on the results, a specification of functional and non-functional requirements was developed, establishing relevant metrics to evaluate user progress, such as the number of solved problems, and user position in a ranking table, among others.
Furthermore, an achievement system was designed, structured into different levels and categories. This system is intended to recognize significant milestones within the platform and reinforce positive learning behaviors. It seeks not only to reward performance but also to encourage consistency, progressive improvement, and exploration.
The application was developed following a modular architecture, ensuring scalability, maintainability, and component reusability. Special attention was given to integration the existing infrastructure of \aer
As a result, a tool that complements the \aer environment has been obtained, providing value to both students and educators.


\section*{Keywords}
\vspace{-0.5em}
\noindent judge, programming, statistics, gamification, visualization, achievements, data, web, React, Node